
### Group project

### State the problem and describe the data set


```{r}
library(car)
```

### Loading Dataset
```{r}
rm(list=ls())
#Load the data set 
#heart_disease <-read.csv("/Users/ishani/Documents/Math 449/heart_2020_cleaned.csv", stringsAsFactors=TRUE) 
heart_disease <- read.csv("/Users/linw/Desktop/449project/heart_2020_cleaned.csv", stringsAsFactors = TRUE)

head(heart_disease) 

#structure and summary of the data 
str(heart_disease) 
summary(heart_disease)
```

### Data cleaning

```{r}
#Outliers

# Save the outlier's index to index_highage
#index_highage <- which(loan_data$age > 122)

# Create data set new_data with outlier deleted
#new_data <- loan_data[-index_highage, ]


### Setting reference variables, ordering levels of ordinal variables
heart_disease$Race = relevel(heart_disease$Race, ref = "White")
heart_disease$GenHealth = factor(heart_disease$GenHealth, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))

# Original linear regression model
glm.fit0=glm(HeartDisease~.,data=heart_disease,family=binomial)
summary(glm.fit0)

attach(heart_disease)

### Creating new quantitative variables for ordinal variables using scores
# AgeScore = c(21,27,32,37,42,47,52,57,62,67,72,77,90)
# DiabeticScore = c(0, 1, 2, 3) # (No, No borderline, Yes, Yes pregnant) 
# GenHeatlhScore = c(0, 1, 2, 3, 4) # (Excellent, Very good, Good, Fair, Poor)
heart_disease$Age=rep(0,319795)
  heart_disease$Age[heart_disease$AgeCategory=="18-24"]=21
  heart_disease$Age[heart_disease$AgeCategory=="25-29"]=27
  heart_disease$Age[heart_disease$AgeCategory=="30-34"]=32
  heart_disease$Age[heart_disease$AgeCategory=="35-39"]=37
  heart_disease$Age[heart_disease$AgeCategory=="40-44"]=42
  heart_disease$Age[heart_disease$AgeCategory=="45-49"]=47
  heart_disease$Age[heart_disease$AgeCategory=="50-54"]=52
  heart_disease$Age[heart_disease$AgeCategory=="55-59"]=57
  heart_disease$Age[heart_disease$AgeCategory=="60-64"]=62
  heart_disease$Age[heart_disease$AgeCategory=="65-69"]=67
  heart_disease$Age[heart_disease$AgeCategory=="70-74"]=72
  heart_disease$Age[heart_disease$AgeCategory=="75-79"]=77
  heart_disease$Age[heart_disease$AgeCategory=="80 or older"]=87
  
heart_disease$DiabeticScore=rep(0,319795)
  heart_disease$DiabeticScore[heart_disease$Diabetic=="No"]=0
  heart_disease$DiabeticScore[heart_disease$Diabetic=="No, borderline diabetes"]=1
  heart_disease$DiabeticScore[heart_disease$Diabetic=="Yes"]=2
  heart_disease$DiabeticScore[heart_disease$Diabetic=="Yes (during pregnancy)"]=3
  
heart_disease$GenHealthScore=rep(0,319795)
  heart_disease$GenHealthScore[heart_disease$GenHealth=="Excellent"]=0
  heart_disease$GenHealthScore[heart_disease$GenHealth=="Very good"]=1
  heart_disease$GenHealthScore[heart_disease$GenHealth=="Good"]=2
  heart_disease$GenHealthScore[heart_disease$GenHealth=="Fair"]=3
  heart_disease$GenHealthScore[heart_disease$GenHealth=="Poor"]=4

# Remove ordinal variables; replaced with scores
heart_disease = subset(heart_disease, select = -c(AgeCategory, Diabetic, GenHealth))
```



### Fit a logistic regression model with all predictors
```{r}
# AgeCategory, Diabetic, GenHealth replaced with scores
glm.fit=glm(HeartDisease ~ ., data=heart_disease, family=binomial)

summary(glm.fit) 
```


### Select the best subset of variables. Perform a diagnostic on the best model.  Perform all possible inferences you can think about.

```{r}
# Model 2: Remove non-significant variables from model with all predictors
# Remove PhysicalActivity
glm.fit2 = update(glm.fit, ~ . -PhysicalActivity) 
summary(glm.fit2)

drop1(glm.fit, test="LRT")
anova(glm.fit2, glm.fit, test="LRT") # likelihood-ratio test comparing models
```

We see that...

```{r}
# Backward elimination
# step(glm.fit)

#Null model
# glm.null = glm(HeartDisease ~ 1, family = binomial, data=heart_disease) 
```


### Use the new model to make predictions.
```{r}

#Confidence Intervals

#Wald confidence interval at alpha=0.05
alpha=0.05
(ci_wald=c(g$coef[2,1]-qnorm(1-alpha/2)* g$coef[2,2], g$coef[2,1]+qnorm(1-alpha/2)* g$coef[2,2]))

# walk CI for the multiplicative effect at alpha=0.05
exp((ci_wald=c(g$coef[2,1]-qnorm(1-alpha/2)* g$coef[2,2], g$coef[2,1]+qnorm(1-alpha/2)* g$coef[2,2]))) 

# LR CI  
# profile likelihood confidence interval
confint(fit) 

# confidence interval for the effect
exp(confint(fit)) 














#Tests

###########################################################

# Wald test

z=g$coef[2,1]/g$coef[2,2]
# just get this stat from the summary
z1=g$coef[2,3]
# notice that z1=z, 

#pval for normal 
p_val=2*(1-pnorm(g$coef[2,3]))
p_val


#p_val for chi^2 distr with df=1
p_val1=1-pchisq(z^2,1)
p_val1

# Both distributions give the same p-value


###########################################################

#Likelihood ratio test
#Model comparison, compare models g(mu)=alpha+beta x  and g(mu)=alpha

# -2 * (L0-L1) , in your summary 
#Null deviance=2(L0-Lsaturated), residual deviance=devuiance= -2*(L1-Lsaturated)
#therefore -2 * (L0-L1)=nul.deviance-deviance

LLstat=(g$null.deviance-g$deviance)
p_val=1-pchisq(LLstat, 1)
p_val


# automated way to perform the likelihood ratio test, Likelihood-ratio test of snoring and heart disease (comparison with the null
drop1(fit, test="Chisq")

#Or can use Anova, but need to install many packages for that
#Anova(fit) 

###########################################################
#Deviance, goodness of fit

#Deviance
Deviance=deviance(fit)
Deviance
# or
Deviance=g$deviance
Deviance

#Goodness of Fit
P_value=1-pchisq(g$deviance, g$df.res)
P_value

################################################################







```


#Use different pi_0 as a cut-off point and create a confusion table.
```{r}
#Chapter 4 Lab-4

glm.pred=rep("Down",1250) # generate a 1250-dim vector with each element as "Down", name this vector glm.pred
glm.pred[glm.probs>.5]="Up" #if the predicted risk > 0.5, assign predicted direction as "up"
table(glm.pred,Direction) # produce the confusion matrix
(507+145)/1250
```


#Perform visualization of data and models.  
```{r}
plot()
pairs()
```

#Plot the ROC curve, find AUC, and the best cutoff point for classification.
```{r}

#install.packages("ROCR")
library(ROCR)

attach(heart_disease)

#install.packages("Epi")
library(Epi)

#Logistic_crabs_all - ROC
hd.roc = ROC(form=HeartDisease~. -AgeCategory -Diabetic -GenHealth -PhysicalActivity, data=heart_disease, plot="ROC")
summary(hd.roc)


ROC(form=(HeartDisease=1)~Age+Asthma,plot="ROC", data=heart_disease, drop.unused.levels = TRUE)

#AUC(glm1.fit)

```


#Perform LOOCV and k-fold cross-validation.

```{r}

#LOOCV - problem 7 ch 5 hw 448
#Logistic_crabs_all - k-fold

#install.packages("DAAG")
library(DAAG)
cv.binary(glm1.fit)


glm.fit3=glm(HeartDisease ~ Age + BMI, data = heart_disease[-1, ], family=binomial)
summary(glm.fit3)

predict.glm(glm.fit3, heart_disease[1, ], type = "response") > 0.5


err <- numeric(nrow(heart_disease))

for (i in 1:nrow(heart_disease)) {
  train <- heart_disease[-i, ]
  #i
  glm.fit4 <- glm(HeartDisease ~ Age + BMI, data = train, family = binomial)
  #ii
  post <- predict(glm.fit4, heart_disease[i, ], type = "response")
  #iii
  pred <- ifelse(post > 0.5, "Yes", "No")
  #iv
  er <- as.integer(pred != heart_disease$HeartDisease[i])
  err[i] <- er
}


err

numOne <- sum(err == 1)

numOne


mean(err)



```




#Try the probit link and the identity links to model data.

```{r}

Malformation.ident=glm(cbind(Present, Absent)~alcohol, family=binomial(link="identity"), data=Malform )
summary(Malformation.ident)


fit1 <- glm(evolved ~ ideology, family=binomial(link="identity"), data=Evo)
summary(fit, dispersion=1)
summary(fit1)


Heart$x <- recode(Heart$snoring, never = 0, occasional = 2, nearly_every_night = 4, every_night = 5)
fit <- glm(yes/(yes+no) ~ x, family=binomial(link=probit),
           + weights=yes+no, data=Heart)
summary(fit)

```


#Which model works better for this data?


#If you have grouped data, use the methods for contingency tables to analyze the data (Chi sq test, G^2, and so on if applicable).

```{r}

#Contingency Table of Data
Malform <- matrix(c(483, 1101, 477, 1121), nrow = 2)

dimnames(Malform)=list(Exercise=c("Regular","Other"),
                       PersonalityType=c("A","B"))
addmargins(Malform)

#Chi Squared Test
chisq.test(Malform)

#G Squared Test
G.test(Malform)

#Person Standardized Residuals
stdres <- chisq.test(Malform)$stdres 
stdres

```
same as midterm question


#Write a report













